---
title: "Pre-Class 2"
author: "Kari Kusler"
date: "9/18/2018"
output: html_document
---
Question 1
```{r setup,include=FALSE, echo=FALSE,eval=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Part 1
#1
exp.draws.1 <- rexp(200)
mean(exp.draws.1)
sd(exp.draws.1)
```
The mean of exp.draws.1 is 0.9153844 and the standard deviation is 0.9082359.

Question 2

```{r echo=FALSE, eval=FALSE}
exp.draws.02 <- rexp(200,0.2)
mean(exp.draws.02)
sd(exp.draws.02)
```
The mean of exp.draws.02 is 4.847144 and the standard deviation is 4.896737.
```{r echo=FALSE, eval=FALSE}
exp.draws.5 <- rexp(200,5)
mean(exp.draws.5)
sd(exp.draws.5)
```
The mean of exp.draws.5 is 0.2079173 and the standard deviation is 0.2017379.
```{r echo=FALSE, eval=FALSE}
exp.draws.7.3 <- rexp(200,7.3)
mean(exp.draws.7.3)
sd(exp.draws.7.3)
```
The mean of exp.draws.7.3 is 0.1224047 and the standard deviation is 0.116555.
```{r echo=FALSE, eval=FALSE}
exp.draws.10 <- rexp(200,10)
mean(exp.draws.10)
sd(exp.draws.10)
```
The mean of exp.draws.10 is 0.115676 and the standard deviation is 0.1238618.

Question 3a

```{r echo=FALSE, eval=TRUE}
hist(exp.draws.1)
```

Question 3b

```{r echo=FALSE, eval=TRUE}
plot(sort(exp.draws.1))
```

Question 3c

```{r echo=FALSE, eval=TRUE}
plot(exp.draws.5,exp.draws.02)
```

```{r echo=FALSE, eval=TRUE}
#Question 4
exp.means <- rep(0,5)
exp.means[1] <- mean(exp.draws.1)
exp.means[2] <- mean(exp.draws.02)
exp.means[3] <- mean(exp.draws.5)
exp.means[4] <- mean(exp.draws.7.3)
exp.means[5] <- mean(exp.draws.10)
exp.rates <- c(1,0.2,5,7.3,10)
exp.sd <- rep(0,5)
exp.sd[1] <- sd(exp.draws.1)
exp.sd[2] <- sd(exp.draws.02)
exp.sd[3] <- sd(exp.draws.5)
exp.sd[4] <- sd(exp.draws.7.3)
exp.sd[5] <- sd(exp.draws.10)
```

Question 4a

```{r echo=FALSE, eval=TRUE}
plot(exp.means,exp.rates)
```

As the mean increases, the rate decreases.

Question 4b

```{r echo=FALSE, eval=TRUE}
plot(exp.sd,exp.rates)
```

As the standard deviation increases, the rate decreases.

Question 4c

```{r echo=FALSE, eval=TRUE}
plot(exp.means,exp.sd)
```

The mean and standard deviation are nearly equal.

Question 5a

```{r echo=FALSE, eval=FALSE}
big.exp.draws.1 <- rexp(1100000)
mean(big.exp.draws.1)
sd(big.exp.draws.1)
```
The mean of big.exp.draws.1 is 0.999491 and the standard deviation is 0.9989396.

Question 5b

```{r echo=FALSE, eval=TRUE}
hist(big.exp.draws.1)
```

Yes it matches (1-e^{-x}), and yes it should because that function was used to create the vector.

Question 5c

```{r echo=FALSE, eval=FALSE}
mean(subset(big.exp.draws.1, big.exp.draws.1>1))
```

The mean of all entries in big.exp.draws.1 greater than 1 is 1.999256.

Question 5d

```{r echo=FALSE, eval=TRUE}
big.exp.draws.1.mat <- matrix(big.exp.draws.1,ncol=1000)
hist(big.exp.draws.1.mat)
big.hist <- hist(big.exp.draws.1.mat)
```

The information necessary to construct the histogram is saved in the histogram object.

Question 5e

```{r echo=FALSE, eval=FALSE}
mean(big.exp.draws.1.mat[,371])
```

The mean of column 371 is 0.9660588.

Question 5f

```{r echo=FALSE, eval=FALSE}
apply(big.exp.draws.1.mat,2,mean)
```
```{r echo=FALSE, eval=TRUE}
hist(apply(big.exp.draws.1.mat,2,mean))
```

The histogram in problem 5b plots a histogram of the distribution, while the histogram in problem 5f plots a histogram of the means of a distribution. The histogram in 5f is approximately normally distributed, as we would expect from the Central Limit Theorem.