# pre-class


Make sure you commit this often with meaningful messages. 



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Read the source code for each of the following three functions, puzzle out what they do, and then brainstorm better names.

```{r}
#checks if the first part of a string matches the prefix that is input
prefix.match <- function(string, prefix) {
  substr(string, 1, nchar(prefix)) == prefix
}
#removes the last element from a list of vectors
remove.last <- function(x) {
  if (length(x) <= 1) return(NULL)
  x[-length(x)]
}
#take a vector x with a certain length and generates a vector with the value y repeated such that the length of the new vector matches the length of the x vector
y.of.length.x <- function(x, y) {
  rep(y, length.out = length(x))
}
```

2. Compare and contrast rnorm() and MASS::mvrnorm(). How could you make them more consistent?

rnorm generates a vector of values taken from a normal distribution and takes the number of values, the mean, and the standard deviation as inputs.
mvrnorm generates multiple vectors of values taken from a multivariate normal distribution and take the number of values, a vector of the means of the variables, a vector specifying the covariance matrix, and additional values as inputs.
They both are ways to generate random variables that follow a normal distribution, in which you can specify the number of samples (n) and the means (mean/mu). However, there are more parameters to specify in the mvrnorm function, and everything except n has a different name. You could make them more consistent by renaming "mu" and "Sigma"" to correspond with "mean"" and "sd" or vice versa.

3. Use `lapply()` and an anonymous function to find the coefficient of variation (the standard deviation divided by the mean) for all columns in the mtcars dataset. 
```{r}
lapply(mtcars, function(x) sd(x)/mean(x))
```

4. Use vapply() to:
    a. Compute the standard deviation of every column in a numeric data frame.
```{r}
v1 <- c(1,3,2,4,5)
v2 <- c(5,4,3,2,1)
data.frame.test <- as.data.frame(rbind(v1,v2))
vapply(data.frame.test,sd,numeric(1))
```
    b. Compute the standard deviation of every numeric column in a mixed data frame. (Hint: you’ll need to use vapply() twice.)
```{r}
#using vapply to filter for only numeric columns
vapply(mtcars[vapply(mtcars,is.numeric,logical(1))],sd,numeric(1))
```